#Assignment. End-to-end data analysis in R. In this assignment, we will use real-world marine data collected as part of Queensland fisheries QFISH database to develop a report (either written on your website or via an R Markdown report embedded in your website) suitable for reporting to your manager, a community group or for your own research.


#Part 1. First, obtain your data.

#ALL DONE
#Open dataset

```{r}

setwd("C:/Users/Danko Castaño Duro/Desktop/Australia 3 semestre/Modulos 1/INTRODUCTION TO R 4/MB5370_Mod04_Danko/MB5370_Mod04_Danko")


shark <- read_csv("shark.csv")

View(shark)
```

#Clean data
#Delete all rows that have the word "total..."
```{r}
library(dplyr)
library(tidyverse)

rows_with_total <- apply(shark, 1, function(row) any(grepl("total", row, ignore.case = TRUE)))

shark_cleaned <- shark[!rows_with_total, ]

view(shark_cleaned)
```
#Explanation
#apply(shark, 1, function(row) ...): This applies the function to each row (MARGIN = 1) of the data frame shark.

#grepl("total", row, ignore.case = TRUE): This checks if the word "total" (case insensitive) is present in any element of the row.

#any(grepl(...)): This returns TRUE if any element in the row contains the word "total".

#rows_with_total: This logical vector indicates which rows contain the word "total".

#shark[!rows_with_total, ]: This subsets the data frame to include only rows that do not contain the word "total".


#Check structure of data set
```{r}
str(shark_cleaned)
```

#Plot data

# Filter the data by Making new tiny datset
```{r}
just_shark_Townsville <- shark_cleaned %>%
  filter(Area == "Townsville", SpeciesGroup == "Shark", CalendarYear >= "2014", CalendarYear <= "2024")
view(just_shark_Townsville)

```

#Plot the dataset 
```{r}
str(just_shark_Townsville)
```


```{r}
ggplot(data = just_shark_Townsville, aes(x = CalendarYear, y = NumberCaught.Total)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(x = "Year", y = "Number of Sharks Caught", title = "Total Number of Sharks Caught in Townsville by Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
```


#Plot for all species in townsvile

#Filter data
```{r}
just_Townsville <- shark_cleaned %>%
  filter(Area =="Townsville", CalendarYear >="2014", CalendarYear <= "2024")

view(just_Townsville)

```

#Plot data

```{r}
ggplot(data = just_Townsville, aes(x = CalendarYear, y = NumberCaught.Total, fill = SpeciesGroup)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  geom_text(aes(label = NumberCaught.Total), position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(x = "Year", y = "Number of Sharks Caught", title = "Total Number of SpeciesGroup Caught in Townsville from 2014-2024") +
  scale_fill_manual(values = c("Shark" = "blue", "Turtle" = "green", "Other" = "white")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
```



