#Assignment. End-to-end data analysis in R. In this assignment, we will use real-world marine data collected as part of Queensland fisheries QFISH database to develop a report (either written on your website or via an R Markdown report embedded in your website) suitable for reporting to your manager, a community group or for your own research.


#Part 1. First, obtain your data.

#ALL DONE

#Part 2. Tidy the data.

```{r}
library(tidyverse)
```

#data successfully download in an "organized way"

#Open dataset

```{r}

setwd("C:/Users/Danko Castaño Duro/Desktop/Australia 3 semestre/Modulos 1/INTRODUCTION TO R 4/MB5370_Mod04_Danko/MB5370_Mod04_Danko")


shark <- read_csv("shark.csv")

View(shark)

```

#Clean data
#Delete all rows that have the word "total..."
```{r}
library(dplyr)

rows_with_total <- apply(shark, 1, function(row) any(grepl("total", row, ignore.case = TRUE)))

shark_cleaned <- shark[!rows_with_total, ]

view(shark_cleaned)
```
#Explanation
#apply(shark, 1, function(row) ...): This applies the function to each row (MARGIN = 1) of the data frame shark.

#grepl("total", row, ignore.case = TRUE): This checks if the word "total" (case insensitive) is present in any element of the row.

#any(grepl(...)): This returns TRUE if any element in the row contains the word "total".

#rows_with_total: This logical vector indicates which rows contain the word "total".

#shark[!rows_with_total, ]: This subsets the data frame to include only rows that do not contain the word "total".


#Check structure of data set
```{r}
str(shark_cleaned)
```


#Plot data

# Filter the data by Making new tiny datset
```{r}
just_shark_Townsville <- shark_cleaned %>%
  filter(Area == "Townsville", SpeciesGroup == "Shark", CalendarYear >= "2014", CalendarYear <= "2024")
view(just_shark_Townsville)


```

#Plot the dataset 



```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)

ggplot(just_shark_Townsville, aes(x = CalendarYear), y = NumberCaught_Total, color = SpeciesGroup) +
  geom_boxplot() +
  labs(title = "Shark Areas by Year Townsville", x = "Year", y = "NumberCaught_Total") +
  theme_minimal()
  

```

library(ggplot2)

ggplot(shark_cleaned, aes(x = as.factor(CalendarYear), y = Area, color = SpeciesGroup)) +
  geom_boxplot() +
  labs(title = "Shark Areas by Year and Species Group", x = "Year", y = "Area") +
  theme_minimal()




